Functional trait data tell new story about biodiversity change

### 2020-10-12 paper/manuscript.Rmd
Flesh out outline

---
title: "manuscript"
author: "Kari Norman"
date: "8/24/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

###Methods
- matched all species to an ID using taxadb, if more than one species name matched to the same ID they were considered the same species and observations of those species were consolidated into one
- only include time series with at least 4 species and at least 4 years
- drop traits that were the same for all observed species
- all non-categorical or binary traits were z-score scaled to give each trait equal weight (@lepsQuantifyingInterpretingFunctional2006, @schleuterUserGuideFunctional2010)
- raw abundance counts within a year for each plot
- needed euclidian distance matrix - used calliez correction because it was a sufficient correction for all plots even though it led to a marginal decrease in "quality" for the corrected hull
- weight by abundance if available and biomass if not
- null model stuff via @swensonBiogeographyFilteringWoody2012
- for years where all species in the species pool were observed the null model sd is zero for functional richness, since FRic does not incorporate abundance changes. 
- drop species with trait match

>>>>

---
title: "Functional trait data tell new story about biodiversity change"
author: 
  - name: "Kari E A Norman"
    affiliation: ucb
  - name: "Carl Boettiger"
    affiliation: ucb, ropensci
address:
  - code: ucb
    address: "Dept of Environmental Science, Policy, and Management, University of California Berkeley, Berkeley CA 94720-3114, USA"
  - code: ropensci
    address: "The rOpenSci Project, University of California Berkeley, Berkeley CA 94720-3114, USA" 
abstract: |

journal: "Ecography"
date: "`r Sys.Date()`"
bibliography: refs.bib
layout: 3p
header-includes:
   - \usepackage{lineno}
   - \linenumbers
output: 
  rticles::elsevier_article:
    includes:
      in_header: preamble.tex
editor_options: 
  markdown: 
    wrap: 72
---

### Introduction

Ecological communities are experiencing unprecedented change as a result
of anthropogenic pressures such as climate change, land use change, and
invasive species.

#### Paragraph 1 - communities are changing, we need to know how and what that means for ecosystem properties

-   communities are changing as a result of a whole bunch of different
    pressures

-   a few well documented phenomena: accelerating extinction rates, what
    else?

-   it's essential for conservation and ecology to understand what these
    changes mean for ecosystem properties like stability, function, etc

#### Paragraph 2 - species diversity is the go to - it's not working

-   traditionally the most common method of capturing change in
    communities is biodiversity

-   despite other well documented changes, species diversity doesn't
    seem to capture local community-level changes
    ([@dornelasAssemblageTimeSeries2014],
    [@suppSpecieslevelCommunitylevelResponses2014],
    [@vellendGlobalMetaanalysisReveals2013]), though there is evidence
    of a significant increase in turnover

-   We need to look beyond species and their abundances to understand
    the kinds of changes happening in communities

#### Paragraph 3 - functional diversity is the answer!

-   Functional diversity provides a potentially powerful tool for
    examining biodiversity changes because it captures characteristics
    beyond species identity.

-   Powerful functional lens ([@mcgillRebuildingCommunityEcology2006])

-   Functional diversity can tells us a story independent of species
    richness, as huge functional changes can occur while species
    richness (or even diversity!) stays the same. This especially gives
    insight into the turnover we know is occurring

-   We examine 100's of mammal, bird, and amphibian timeseries from the
    Biotime database paired with functional trait data to assess overall
    trends in functional change and identify where functional diversity
    tells us a story beyond species richness

### Methods

#### *Data*

Biodiversity timeseries were obtained from the Biotime database, a
global repository of assemblage timeseries with 386 individual studies
from multiple taxa.

-   matched all species to an ID using taxadb, if more than one species
    name matched to the same ID they were considered the same species
    and observations of those species were consolidated into one - only
    include time series with at least 4 species and at least 4 years

-   drop traits that were the same for all observed species

-   all non-categorical or binary traits were z-score scaled to give
    each trait equal weight
    ([@lepsQuantifyingInterpretingFunctional2006],
    [@schleuterUserGuideFunctional2010])

-   raw abundance counts within a year for each plot

-   needed euclidian distance matrix - used calliez correction because
    it was a sufficient correction for all plots even though it led to a
    marginal decrease in "quality" for the corrected hull

-   weight by abundance if available and biomass if not

-   null model stuff via [@swensonBiogeographyFilteringWoody2012] - for
    years where all species in the species pool were observed the null
    model sd is zero for functional richness, since FRic does not
    incorporate abundance changes. - drop species with trait match
